In this section, we recall and develop some theory of (quasi)modular forms.

\subsection{Modular forms and examples}

Let $\h$ be the upper half-plane $\{z\in\C\mid\Im(z)>0\}$.
\begin{lemma}\label{def:Gamma-1-Action}\lean{UpperHalfPlane.SLAction}\leanok
    The modular group $\Gamma_1:=\mathrm{SL}_2(\Z)$ acts on $\h$ by linear fractional transformations
$$\left(\begin{smallmatrix}a&b\\c&d\end{smallmatrix}\right)z:=\frac{az+b}{cz+d}.$$
\end{lemma}
\begin{proof}\leanok
\end{proof}

Let $N$ be a positive integer.
\begin{definition}\label{def:level-N-princ-cong-subgp}\lean{CongruenceSubgroup.Gamma}\leanok
The \emph{level $N$ principal congruence subgroup} of $\Gamma_1$ is
$$
    \Gamma(N):=\left\{\left.\left(\begin{smallmatrix}a&b\\c&d\end{smallmatrix}\right)\in\Gamma_1\right|\left(\begin{smallmatrix}a&b\\c&d\end{smallmatrix}\right)\equiv\left(\begin{smallmatrix}1&0\\0&1\end{smallmatrix}\right)\;\mathrm{mod}\;N\right\}.
$$
\end{definition}

\begin{definition}\label{def:congruence-subgroup}\uses{def:level-N-princ-cong-subgp}\lean{CongruenceSubgroup.IsCongruenceSubgroup}\leanok
    A subgroup $\Gamma\subset\Gamma_1$ is called a \emph{congruence subgroup} if $\Gamma(N)\subset\Gamma$ for some $N\in\N$.
\end{definition}

\begin{definition}\label{def:Gamma-generators}\uses{def:level-N-princ-cong-subgp}\lean{ModularGroup.S,ModularGroup.T,α,β}\leanok
  Define the matrices

  \[
    S = \begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix} \in \Gamma_1,
    T = \begin{pmatrix} 1 & 1 \\ 0 & 1 \end{pmatrix} \in \Gamma_1,
    \alpha = \begin{pmatrix} 1 & 2 \\ 0 & 1 \end{pmatrix} \in \Gamma_2 \subset \Gamma_1,
    \beta = \begin{pmatrix} 1 & 0 \\ 2 & 1 \end{pmatrix} \in \Gamma_2 \subset \Gamma_1.
  \]

  It is easily verifiable that $\alpha = T^2$ and $\beta = -S\alpha^{-1}S = -ST^{-2}S$.
\end{definition}

The following two lemmas tell us the group structure of $\Gamma(1) = \Gamma_1$ and $\Gamma(2)$, which we will use later on to define the theta forms.

\begin{lemma}\label{lemma:Gamma-1-generators}\uses{def:Gamma-generators}\lean{SL2Z_generate}\leanok
  We have $\Gamma(1) = \langle S, T, -I \rangle$.
\end{lemma}
\begin{proof}\leanok
  See~\cite[Exercise 1.1.1]{first course}.
\end{proof}

\begin{lemma}\label{lemma:Gamma-2-generators}\uses{def:Gamma-generators}\lean{Γ2_generate}\leanok
  We have $\Gamma(2) = \langle \alpha, \beta, -I \rangle$.
\end{lemma}
\begin{proof}\leanok
    See~\cite[Exercise 1.2.4]{first course}.
\end{proof}

Let $z\in\h$, $k\in\Z$, and $\left(\begin{smallmatrix}a&b\\c&d\end{smallmatrix}\right)\in\mathrm{SL}_2(\Z)$. We omit many of the proofs below when they exist in Mathlib already.
\begin{definition}\label{def:automorphy-factor}\leanok \lean{UpperHalfPlane.denom}
    The \emph{automorphy factor} of weight $k$ is defined as
$$j_k(z,\left(\begin{smallmatrix}a&b\\c&d\end{smallmatrix}\right)):=(cz+d)^{-k}.$$
\end{definition}

\begin{lemma}\label{lemma:automorphy-factor-chain-rule}\uses{def:automorphy-factor}\leanok \lean{UpperHalfPlane.denom_cocycle}
    The automorphy factor satisfies the \emph{chain rule}
$$j_k(z,\gamma_1\gamma_2)=j_k(z,\gamma_1)\,j_k(\gamma_2z,\gamma_1). $$
\end{lemma}
\begin{proof}\leanok
\end{proof}

\begin{definition}\label{def:slash-operator}\uses{def:automorphy-factor, def:Gamma-1-Action}\lean{ModularForm.SL_slash_apply}\leanok
    Let $F$ be a function on $\h$ and $\gamma\in\mathrm{SL}_2(\Z)$. Then the \emph{slash operator} acts on $F$ by
$$(F|_k\gamma)(z):=j_k(z,\gamma)\,F(\gamma z). $$
\end{definition}

\begin{lemma}\label{lemma:slash-operator-chain-rule}\uses{lemma:automorphy-factor-chain-rule}\lean{SlashAction.slash_mul}\leanok
  The chain rule implies $$F|_k\gamma_1\gamma_2=(F|_k\gamma_1)|_k\gamma_2.$$
\end{lemma}
\begin{proof}\leanok
\end{proof}

In particular, this lemma implies that if $\Gamma = \langle M_i \rangle_{i \in \mathcal{I}} \rangle$, then the slash action $F|\gamma$ is uniquely determined by the action of generators, i.e. $F|M_i$ and $F|M_i^{-1}$.

\begin{lemma}\label{lemma:slash-negI-even-weight}\uses{def:slash-operator}\lean{modular_slash_negI_of_even}\leanok
   For even $k$, $F|_{k}(-I) = F$.
\end{lemma}
\begin{proof}\leanok
Follows from the definition of the slash operator:
$(F|_{k}(-I))(z) = (-1)^{-k}F((-I)z) = F(z)$.
\end{proof}

\begin{definition}\label{def:holomorphic-modular-form}\uses{def:congruence-subgroup}\lean{ModularForm}\leanok
	Let  $\Gamma$ denote a subgroup of $\mathrm{SL}_2(\mathbb{Z})$, then a modular form  of level $\Gamma$ and weight $k \in \mathbb{Z}$ is a function $f : \mathbb{H} \to \mathbb{C}$ such that:
	\begin{enumerate}
		\item For all $\gamma \in \Gamma$ we have $f\mid_k \gamma = f$  (such functions are called slash invariant).
		\item $f$ is holomorphic on $\mathbb{H}$.
		\item For all $\gamma \in \mathrm{SL}_2(\mathbb{Z})$, there exist $A, B \in \mathbb{R}$ such that for all $z \in \mathbb{H}$, with $ A \le \mathrm{Im}(z)$, we have $|(f \mid_k \gamma) (z) |\le B$. Here $| - |$ denotes the standard complex absolute value.
	\end{enumerate}
\end{definition}
This defines a complex vector space which we denote by $M_{k}(\Gamma)$. By replacing condition $(3)$ above with (4) below defines the subspace of cusp forms, which we denote by $S_k(\Gamma)$.

\begin{enumerate}
	 \setcounter{enumi}{3}
	\item  For all $\gamma \in \mathrm{SL}_2(\mathbb{Z})$, and all  $0 < \epsilon$, there exists $A \in \mathbb{R}$ such that for all $z \in \mathbb{H}$, with $ A \le \mathrm{Im}(z)$, we have $|(f \mid_k \gamma) (z) |\le \epsilon$.
\end{enumerate}


\begin{definition}\label{def:Mk}\uses{def:holomorphic-modular-form}\lean{ModularForm}\leanok
    Let $M_k(\Gamma)$ be the space of modular forms of weight $k$ and congruence subgroup $\Gamma$.
\end{definition}


Let us consider several examples of modular forms.
\begin{definition}\label{def:Ek} \lean{ModularForm.eisensteinSeries_MF }\leanok
For an even integer $k\geq 4$ we define the \emph{weight $k$ Eisenstein series} as
\begin{equation}\label{eqn:Ek-definition}
E_k(z):=\frac{1}{2}\sum_{(c,d)\in\Z^2, (c,d)=1}(cz+d)^{-k}.\end{equation}
\end{definition}
\begin{lemma}\label{lemma:Ek-is-modular-form}\uses{def:Mk, def:Ek}\lean{EisensteinSeries.eisensteinSeries_SIF}\leanok
For all $k$, $E_k\in M_k(\Gamma_1)$.
Especially, we have
\begin{equation}\label{eqn:Ek-trans-S}
    E_k \left(-\frac{1}{z}\right) = z^k E_k(z).
\end{equation}
\end{lemma}
\begin{proof}\leanok
This follows from the fact that the sum converges absolutely.
Now apply slash operator with $\gamma = \left(\begin{smallmatrix} 0 & -1 \\ 1 & 0 \end{smallmatrix}\right)$ gives \eqref{eqn:Ek-trans-S}.
\end{proof}

Since modular forms are holomorphic and periodic they have Fourier expansions, and their Fourier coefficients have polynomial growth. We use the corresponding results from Mathlib in the formalization.

For Eisenstein series, we can see this directly from the following:

\begin{lemma}\label{lemma:Ek-Fourier}\uses{def:Ek}\lean{E_k_q_expansion}\leanok
The Eisenstein series possesses the Fourier expansion
\begin{equation}\label{eqn:Ek-Fourier}E_k(z)=1+\frac{2}{\zeta(1-k)}\sum_{n=1}^\infty \sigma_{k-1}(n)\,e^{2\pi i z}, \end{equation}
where $\sigma_{k-1}(n)\,=\,\sum_{d|n} d^{k-1}$. In particular, we have
\begin{align}
  E_4(z)\,=\,& 1+240\sum_{n=1}^\infty \sigma_3(n)\,e^{2\pi i n z} \notag \\
  E_6(z)\,=\,& 1-504\sum_{n=1}^\infty \sigma_5(n)\,e^{2\pi i n z}. \notag
\end{align}
\end{lemma}
\begin{proof}\leanok
\end{proof}

The infinite sum \eqref{eqn:Ek-definition} does not converge absolutely for $k=2$.
On the other hand, the expression \eqref{eqn:Ek-Fourier} converges to a holomorphic function on the upper half-plane and we will take it as a definition of $E_2$ (See Definition \ref{def:E2} below).


The discriminant form is a unique normalized cusp form of weight 12, which can be defined as:
\begin{definition}\label{def:disc-definition}\lean{Delta}\leanok\uses{def:dedekind_eta}
The \emph{discriminant form} $\Delta(z)$ is given by
\begin{equation}\label{eqn:disc-definition}
\Delta(z) = e^{2 \pi i z} \prod_{n \ge 1} (1 - e^{2 \pi i n z})^{24}.
\end{equation}
\end{definition}

This product formula allows us to prove positivity of $\Delta(it)$ for $t > 0$ later. But we need to first check its a modular form. For this we first need some definitions/ results.

We define it as a $q$-series, which gives a holomorphic function on $\mathfrak{H}$.
\begin{definition}\label{def:E2}\lean{E₂_eq}\leanok
We set
\begin{equation}\label{eqn:E2}
    E_2(z):= 1-24\sum_{n=1}^\infty \sigma_1(n)\,e^{2\pi i n z}.
\end{equation}
\end{definition}

\begin{lemma}\label{lemma:E2-transform-S}\uses{def:E2}\leanok\lean{E₂_transform}
This function is not modular, however it satisfies
\begin{equation}\label{eqn:E2-S-transform}
    z^{-2}\,E_2\left(-\frac{1}{z}\right) = E_2(z) -\frac{6i}{\pi}\, \frac{1}{z}.
\end{equation}
\end{lemma}

\begin{proof}\leanok
This is exercise 1.2.8 of \cite{first course}.
\end{proof}

More generally, we have
\begin{lemma}\label{lemma:E2-transform-general}\leanok\lean{G₂_transform}\uses{lemma:E2-transform-S, def:E2}
\begin{equation}\label{eqn:E2-transform-general}
(cz + d)^{-2} E_2\left(\frac{az + b}{cx + d}\right) = E_2(z) - \frac{6ic}{\pi (cz + d)}, \quad \begin{pmatrix} a & b \\ c & d\end{pmatrix} \in \mathrm{SL}_{2}(\mathbb{Z}).
\end{equation}
\end{lemma}
\begin{proof}\leanok
    Use the fact that $\mathrm{SL}_2(\mathbb{Z})$ is generated by $S$ and $T$.
    Then \eqref{eqn:E2-transform-general} follows from \eqref{eqn:E2-S-transform} and $E_2|_T = E_2$.
\end{proof}

\begin{definition}\label{def:dedekind_eta}\leanok \lean{η}
The Dedekind eta function is defined as
$$
    \eta(z) = q^{1/24} \prod_{n \ge 1} (1 - q^n)
$$
where $q = e^{2\pi i z}$.
\end{definition}

\begin{lemma}\label{lemma:dedekind_eta_transformation}\lean{eta_equality}\uses{def:dedekind_eta}\leanok
The Dedekind eta function transforms as
$$
    \eta\left(-\frac{1}{z}\right) = \sqrt{-iz} \eta(z).
$$
\end{lemma}
\begin{proof}\leanok
    Consider the logarithmic derivative of $\eta$, which one can easily see is equal to $\frac{\pi i}{12} E_2$.
    The result then follows from the transformation of $E_2$.

    See \cite[proposition 1.2.5]{first course}.
\end{proof}


\begin{lemma}\label{lemma:disc-cuspform}\uses{def:disc-definition, lemma:dedekind_eta_transformation}\leanok\lean{Delta}
$\Delta(z) \in M_{12}(\Gamma_1)$.
Especially, we have
\begin{equation}\label{eqn:disc-trans-S}
    \Delta\left(-\frac{1}{z}\right) = z^{12} \Delta(z).
\end{equation}
Also, it vanishes at the unique cusp, i.e. it is a cusp form of level $\Gamma_1$ and weight $12$.
\end{lemma}
\begin{proof}\leanok
    The fact that it is invariant under translation is clear from the definition, so we only need to check transformation under $S$. Now, note that $\eta^{24} = \Delta$, and from \ref{lemma:dedekind_eta_transformation} we have $\eta(-1/z) = \sqrt{-iz} \eta(z)$, so $\Delta(-1/z) = z^{12} \Delta(z)$ as required.
\end{proof}



\begin{lemma}\label{lemma:disc-E4E6}\uses{def:disc-definition}\lean{Delta_E4_eqn}\leanok
We have
\begin{equation}
\Delta(z) = (E_4^3-E_6^2)/1728.
\end{equation}
\end{lemma}
\begin{proof}\leanok
    We only need to show its a cuspform, since once we have this, dividing the rhs by $\Delta$ would give a modular form of weight $0$ which is a constant, and so we can determine the constant easily.

    To check its a cuspform, we just look at  the $q$-expansions of $E_4$ and $E_6$ and prove directly that the first term vanishes.
\end{proof}

%Note that the RHS of \eqref{eqn:disc-logder} is equal to the $E_2(z)$.
%As a side note, we can also consider defining $\Delta$ as \eqref{eqn:disc-prodformula}, and prove that it coincides with \eqref{eqn:disc-definition}.
%Such an argument can be found in \cite[Section 2.4]{Bruinier}.

\begin{corollary}\label{cor:disc-pos}\uses{def:disc-definition}\lean{Delta_imag_axis_pos}\leanok
$\Delta(it) > 0$ for all $t > 0$.
\end{corollary}
\begin{proof}\leanok
By \ref{def:disc-definition}, we have
$$
\Delta(it) = e^{-2 \pi t} \prod_{n \ge 1} (1 - e^{-2 \pi n t})^{24} > 0.
$$
\end{proof}

The following nonvanishing result, which directly follows from \Cref{def:disc-definition}, will be used in the construction of the magic function.
\begin{corollary}\label{cor:disc-nonvanishing}\uses{def:disc-definition}\leanok\lean{Δ_ne_zero}
$\Delta(z) \neq 0$ for all $z \in \h$.
\end{corollary}
\begin{proof}\leanok
This follows from the product formula.
\end{proof}


A key fact in the theory of modular forms is that the spaces $M_k(\Gamma)$ are finite-dimensional.
To prove this we will do use the following non-standard proof. First we have the following result.

\begin{theorem}\label{thm:nonpos_wt}\uses{def:holomorphic-modular-form}\lean{ModularForm}\leanok
    Let $k \in \Z$ with $k < 0$. Then $M_k(\Gamma_1) = \{0\}$ and moreover $\dim M_0(\Gamma(1)) = 1$.
\end{theorem}
\begin{proof}\leanok
    The proof makes use of the maximum modulus principle, as its already been formalised we skip the details here but see the lean proof for details.
\end{proof}


\begin{theorem}\label{thm:lvl1_dims}\uses{def:holomorphic-modular-form, lemma:disc-E4E6, def:disc-definition}\lean{ModularForm.dimension_level_one}\leanok
    Let $k \in \Z$ with $k \ge 0$ and even. Then $\dim M_k(\Gamma_1) = \lfloor k / 12 \rfloor $ if $k \equiv 2 \mod 12$ and $\dim M_k(\Gamma_1) = \lfloor k / 12 \rfloor + 1$ if $k \not\equiv 2 \mod 12$.
\end{theorem}
\begin{proof}\leanok
First we note that for $2 < k$ we have $\dim(M_k(\Gamma_1)) = 1 + \dim S_k(\Gamma_1)$. This follows since we know the $E_k$ are in $M_k$ so by scaling appropriately, any non-cuspform $f \in M_k$ we would have $f - a E_k \in S_k$ for some $a$.

Next, note that  $S_k(\Gamma_1)$ is isomorphic to $M_{k-12}(\Gamma_1)$, since if $f \in S_k$ then $f/ \Delta$ is now a modular form (using the product expansion of $\Delta$ and its non-vanishing on $\mathfrak{H}$) of weight $k-12$. Note its important that $f$ is a cuspform so that the quotient by $\Delta$ is a modular form.

So we only need to know the dimensions of $M_k(\Gamma_1)$ for $0 \le k \le 12$. For $k = 0$ we have $\dim M_0(\Gamma_1) = 1$ by \Cref{thm:nonpos_wt}.  For $k = 4$ we have $\dim M_4(\Gamma_1) = 1$ since if there was a cuspform $f$ of weight $4$ then $f/ \Delta$ would be a modular form of negative weight, i.e. zero, so $f=0$. Similarly for $k=6,8,10$. For $k=12$ we have $\dim S_{12}(\Gamma_1) = 1$ since the discriminant form is a cusp form of weight $12$ and any other cusp form of weight $12$ would be a scalar multiple of $\Delta$ (since their ratio would be a modular form of weight $0$). So we have $\dim M_{12}(\Gamma_1) = 2$.

Finally we need to check that $\dim M_2(\Gamma_1) = 0$. Firstly, there can't be any cuspforms here by the same argument as above. So we need to check that there are no modular forms of weight $2$. If we did have one, call it $f$ then $f^2$ would be a non-cuspform of weight $4$ and so $f^2 = a E_4$, where in fact $a=a_0(f)^2$ (since $(f^2-a_0(f)E_4)$ is now a cuspform of weight $4$ which means its zero). Similarly, $f^3 = a_0(f)^3 E_6$. But now taking powers to make them weight $12$ forms we see that $a_0(f)^6(E_4^3 - E_6^2) = 0 = 1728 a_0(f)^6 \Delta$
but $a_0(f) \ne 0$ (since its assumed to not be a cuspform), this would mean $\Delta =0$ which we know can't happen.
\end{proof}

\begin{theorem}\label{thm:dim-mf-general-level}\uses{def:Mk, thm:lvl1_dims}\lean{dim_gen_cong_levels}\leanok
Let $\Gamma$ be a congruence subgroup. Then $M_k(\Gamma)$ is finite-dimensional.
\end{theorem}
\begin{proof}\leanok
We know that $\dim(M_k(\Gamma_1))$ is finite dimensional from the above, now this means that there is some $r_k$ such that any element of $M_k(\Gamma_1)$ vanishing at infinity to degree $> r_k$ must be zero. Now take $f \in M_k(\Gamma)$ and vanishes to degree $n$ at infinity, then consider $F = \prod_\gamma f\mid_k \gamma$ where the product is over a set of representatives of $\Gamma_1 \backslash \Gamma$. Then $F$ is a modular form of weight $k d$ where $d = [\Gamma_1: \Gamma]$ and vanishes at infinity to degree at least $n$. So if $n > r_{kd}$ then $F=0$ meaning the $f=0$.
\end{proof}

% the following theorem:
% \begin{theorem}\label{theorem-Mk-finite-dimensional}\uses{def:Mk}
%     The spaces $M_k(\Gamma)$ are finite dimensional.
% \end{theorem}
% \begin{proof}
%   In this project, we only require the theorem for $\Gamma = \Gamma_1$ and $\Gamma = \Gamma_2$. For the proof, see~\cite{Serre73}.
% \end{proof}
% A key fact in the theory of modular forms is that they form a finite-dimensional vector space (for a fixed weight and level).
% Especially, we use the following dimension results.

\begin{corollary}\label{cor:dim-mf}\uses{def:Mk, thm:lvl1_dims}\lean{ModularForm.dimension_level_one}\leanok
We have
\begin{align}
    \dim M_2(\mathrm{SL}_{2}(\mathbb{Z})) &= 0, \label{eqn:dimM2} \\
    \dim M_4(\mathrm{SL}_{2}(\mathbb{Z})) &= 1, \label{eqn:dimM4} \\
    \dim M_6(\mathrm{SL}_{2}(\mathbb{Z})) &= 1, \label{eqn:dimM6} \\
    \dim M_8(\mathrm{SL}_{2}(\mathbb{Z})) &= 1, \label{eqn:dimM8} \\
    \dim S_4(\mathrm{SL}_{2}(\mathbb{Z})) &= 0, \label{eqn:dimS4} \\
    \dim S_6(\mathrm{SL}_{2}(\mathbb{Z})) &= 0, \label{eqn:dimS6} \\
    \dim S_8(\mathrm{SL}_{2}(\mathbb{Z})) &= 0. \label{eqn:dimS8}
\end{align}
\end{corollary}
\begin{proof}\leanok
\end{proof}

Another examples of modular forms we would like to consider are \emph{theta functions} \cite[Section~3.1]{1-2-3}.
\begin{definition}\label{def:th00-th01-th10}\lean{Θ₂, Θ₃, Θ₄}\leanok
We define three different theta functions (so called ``Thetanullwerte'') as
\begin{align}
  \Theta_{2}(z) = \theta_{10}(z)\,=\, & \sum_{n\in\Z}e^{\pi i (n+\frac12)^2 z}. \notag \\
  \Theta_{3}(z) = \theta_{00}(z)\,=\, & \sum_{n\in\Z}e^{\pi i n^2 z} \notag \\
  \Theta_{4}(z) = \theta_{01}(z)\,=\, & \sum_{n\in\Z}(-1)^n\,e^{\pi i n^2 z} \notag \\
\end{align}
\end{definition}

For convenience, we use the following notations for the fourth powers of the theta functions.
\begin{definition}\label{def:H2-H3-H4}\uses{def:th00-th01-th10}\lean{H₂, H₃, H₄}\leanok
Define
\begin{equation}
    H_2 = \Theta_2^4, \quad H_3 = \Theta_3^4, \quad H_4 = \Theta_4^4. \label{eqn:H2-H3-H4}
\end{equation}
\end{definition}
Note that we only need these fourth powers to define \eqref{def:b-definition}.

The group $\Gamma_1$ is generated by the elements $T=\left(\begin{smallmatrix}1&1\\0&1\end{smallmatrix}\right)$, $S=\left(\begin{smallmatrix}0&1\\-1&0\end{smallmatrix}\right)$, and $-I = \left(\begin{smallmatrix}-1&0\\0&-1\end{smallmatrix}\right)$ (\Cref{lemma:Gamma-1-generators}), and the transformation of functions under $\Gamma(2)$ is determined by that under $\Gamma_1$ (by \Cref{lemma:slash-operator-chain-rule}).
The following lemma shows how the theta functions (and their powers) transform under the slash action of these matrices.

\begin{lemma}\label{lemma:theta-transform-S-T}\uses{def:th00-th01-th10, def:H2-H3-H4}\lean{H₂_T_action,H₃_T_action,H₄_T_action,H₂_S_action,H₃_S_action,H₄_S_action}\leanok
These elements act on the theta functions in the following way
\begin{align}
    H_2 | S &= -H_4 \label{eqn:H2-transform-S} \\
    H_3 | S &= -H_3 \label{eqn:H3-transform-S} \\
    H_4 | S &= -H_2 \label{eqn:H4-transform-S}
\end{align}
and
\begin{align}
    H_2 | T &= -H_2 \label{eqn:H2-transform-T} \\
    H_3 | T &= H_4 \label{eqn:H3-transform-T} \\
    H_4 | T &= H_3 \label{eqn:H4-transform-T}
\end{align}
\end{lemma}
\begin{proof}\leanok
The last three identities easily follow from the definition.
For example, \eqref{eqn:H2-transform-T} follows from
\begin{align}
    \Theta_{2}(z + 1) &= \sum_{n\in\Z}e^{\pi i (n+\frac12)^2 (z + 1)}
    = \sum_{n \in \Z} e^{\pi i (n + \frac{1}{2})^{2}} e^{\pi i (n + \frac{1}{2})^{2} z} \\
    &= \sum_{n \in \Z} e^{\pi i (n^2 + n + \frac{1}{4})} e^{\pi i (n + \frac{1}{2})^{2} z} = \sum_{n \in \Z} (-1)^{n^2 + n}e^{\pi i / 4} e^{\pi i (n + \frac{1}{2})^{2} z} \\
    &= e^{\pi i / 4} \Theta_{2}(z)
\end{align}
and taking 4th power.
\eqref{eqn:H2-transform-S} and \eqref{eqn:H4-transform-S} are equivalent under $z \leftrightarrow -1/z$, so it is enough to show \eqref{eqn:H2-transform-S} and \eqref{eqn:H3-transform-S}.
These identities follow from the identities of the \emph{two-variable} Jacobi theta function, which is defined as (be careful for the variables, where we use $\tau$ instead of $z$)
\begin{equation}
    \theta(z, \tau) = \sum_{n \in \mathbb{Z}} e^{2 \pi i n z + \pi i n^2 \tau} \label{eqn:jacobi2}
\end{equation}
and already formalized by David Loeffler.
This function specialize to the theta functions as
\begin{align}
    \Theta_{2}(\tau) &= e^{\pi i \tau / 4} \theta(-\tau / 2, \tau) \label{eqn:Th2-as-jacobi2} \\
    \Theta_{3}(\tau) &= \theta(0, \tau) \label{eqn:Th3-as-jacobi2} \\
    \Theta_{4}(\tau) &= \theta(1/2, \tau) \label{eqn:Th4-as-jacobi2} \\
\end{align}

Poisson summation formula gives
\begin{equation}
    \theta(z, \tau) = \frac{1}{\sqrt{-i \tau}} e^{-\frac{\pi i z^2}{\tau}} \theta\left(\frac{z}{\tau}, -\frac{1}{\tau}\right) \label{eqn:jacobi2transform}
\end{equation}
and applying the specializations above yield the identities.
For example, \eqref{eqn:H4-transform-S} follows from
\begin{equation}
    \Theta_{4}(\tau) = \theta\left(\frac{1}{2}, \tau\right) = \frac{1}{\sqrt{-i\tau}} e^{- \frac{\pi i }{4 \tau}} \theta\left(\frac{1}{2 \tau}, -\frac{1}{\tau}\right) = \frac{1}{\sqrt{-i\tau}} \Theta_{2}\left(-\frac{1}{\tau}\right)
\end{equation}
and taking 4th power.
\end{proof}

Using the above identities, we can prove that these are modular forms.
\begin{lemma}\label{lemma:theta-slash-invariant}\uses{lemma:slash-operator-chain-rule,lemma:slash-negI-even-weight,lemma:theta-transform-S-T,lemma:Gamma-2-generators}\lean{H₂_SIF,H₃_SIF,H₄_SIF}\leanok
  $H_{2}$, $H_{3}$, and $H_{4}$ are slash invariant under $\Gamma(2)$, i.e. for all $\gamma \in \Gamma(2)$ and $i \in \{2, 3, 4\}$, we have $H_i|\gamma = H_i|\gamma^{-1} = H_i$.
\end{lemma}
\begin{proof}\leanok
  By \cref{lemma:Gamma-2-generators} and \cref{lemma:slash-operator-chain-rule}, it suffices to show that the $H_i$ are invariant under slash actions with respect to $\alpha$, $\beta$, and $-I$.
Invariance under $-I$ follows from Lemma \ref{lemma:slash-negI-even-weight}.
The rest follows from Lemma \ref{lemma:slash-operator-chain-rule}, \ref{lemma:theta-transform-S-T}, and the matrix identities
\begin{equation}
    \alpha = T^2, \quad \beta = -S\alpha^{-1}S = -ST^{-2}S. \label{eqn:matrix}
\end{equation}
For example, invariance for $H_2$ can be proved by
\begin{align}
    H_2|\alpha &= H_2 |T^{2} = -H_2 |T = H_2 \\
    H_2|\beta &= H_2 |(-S\alpha^{-1}S) = H_2 | (S\alpha^{-1}S) =-H_4 |(\alpha^{-1}S) = -H_4 |S  = H_2.
\end{align}
\end{proof}

\begin{lemma}\label{lemma:theta-bounded-im-infty}\uses{lemma:theta-slash-invariant, lemma:Gamma-1-generators}\lean{isBoundedAtImInfty_H_slash}\leanok
  For all $\gamma \in \Gamma_1$, $H_{2}|_2 \gamma$, $H_{3}|_2 \gamma$, and $H_{4}|_2 \gamma$ are holomorphic at $i\infty$.
\end{lemma}
\begin{proof}\leanok
    We want to show that for $\gamma \in \Gamma_1$, $\|H_2|_2\gamma(z)\|$ is bounded as $z \in \mathbb{H} \to i\infty$. Firstly, by \Cref{lemma:theta-transform-S-T}, \Cref{lemma:Gamma-2-generators} and induction on group elements, we notice that $\{\pm H_2, \pm H_3, \pm H_4\}$ is closed under action by $\Gamma_1$. Hence, it suffices to prove that $H_2$, $H_3$ and $H_4$ are bounded at $i\infty$. Consider $z \in \mathbb{H}$ with $\Im(z) \geq A$. We proceed by direct algebraic manipulation:
    \begin{align}
        \|H_2(z)\|
        &= \left\|\sum_{n \in \Z} \exp\left(\pi i \left(n + \frac{1}{2}\right)^2 z\right)\right\|^4
        \leq \left(\sum_{n \in \Z} \left\|\exp\left(\pi i \left(n + \frac{1}{2}\right)^2 z\right)\right)\right\|^4 \\
        &= \left(\sum_{n \in \Z} \left\|\exp\left(-\pi \left(n + \frac{1}{2}\right)^2 \Im(z)\right)\right)\right\|^4
        \leq \left(\sum_{n \in \Z} \left\|\exp\left(-\pi \left(n + \frac{1}{2}\right)^2 A\right)\right)\right\|^4
    \end{align}

    Where we prove the final term is convergent by noticing that it equals $\exp(-\pi A / 4)\theta(iA / 2, iA)$, which has been shown to converge in \texttt{Mathlib}. The proofs for $H_3$ and $H_4$ are similar (actually easier) and have been omitted.
\end{proof}

\begin{lemma}\label{lemma:theta-modular}\uses{lemma:theta-slash-invariant,lemma:theta-bounded-im-infty}\lean{H₂_MF,H₃_MF,H₄_MF}\leanok
$H_{2}$, $H_{3}$, and $H_{4}$ belong to $M_2(\Gamma(2))$.
\end{lemma}
\begin{proof}\leanok
\end{proof}


They have Fourier expansions as follows.
\begin{proposition}\label{prop:H2-fourier}\uses{def:H2-H3-H4}\lean{H₂}\leanok
$H_2$ admits a Fourier series of the form
\begin{equation}
    H_2(z) = \sum_{n \ge 1} c_{H_2}(n) e^{\pi i n z}
\end{equation}
for some $c_{H_2}(n) \in \R_{\ge 0}$, with $c_{H_2}(1) = 16$ and $c_{H_2}(n) = O(n^k)$ for some $k \in \N$.
\end{proposition}
\begin{proof}\leanok
We have
\begin{align}
    H_2(z) &= \Theta_2(z)^4 \\
    &= \left(\sum_{n \in \Z} e^{\pi i (n + \frac{1}{2})^{2} z}\right)^{4} \\
    &= \left(2\sum_{n \ge 0} e^{\pi i (n + \frac{1}{2})^{2} z}\right)^{4} \\
    &= \left(2 e^{\pi i z / 4} + 2 \sum_{n \ge 1} e^{\pi i (n^2 + n + \frac{1}{4}) z}\right)^{4} \\
    &= 16 e^{\pi i z}\left(1 + \sum_{n \ge 1} e^{\pi i (n^2 + n)z}\right)^{4} \\
    &= 16 e^{\pi i z} + \sum_{n \ge 2} c_{H_2}(n) e^{\pi i n z} \\
    &= \sum_{n \ge 1} c_{H_2}(n) e^{\pi i n z}.
\end{align}
\end{proof}

\begin{proposition}\label{prop:H3-fourier}\uses{def:H2-H3-H4}\lean{H₃}\leanok
$H_3$ admits a Fourier series of the form
\begin{equation}
    H_3(z) = \sum_{n \ge 0} c_{H_3}(n) e^{\pi i n z}
\end{equation}
for some $c_{H_3}(n) \in \R_{\ge 0}$ with $c_{H_3}(0) = 1$ and $c_{H_3}(n) = O(n^k)$ for some $k \in \N$.
Especially, $H_3$ is not cuspidal.
\end{proposition}
\begin{proof}\leanok
We have
\begin{equation}
    H_3(z) = \Theta_3(z)^{4} = \left(\sum_{n \in \Z} e^{\pi i n^2 z}\right)^{4} = \left(1 + 2 \sum_{n \ge 1} e^{\pi i n^2 z}\right)^{4} = 1 + O(e^{\pi i z}).
\end{equation}
\end{proof}

\begin{proposition}\label{prop:H4-fourier}\uses{def:H2-H3-H4}\lean{H₄}\leanok
$H_4$ admits a Fourier series of the form
\begin{equation}
    H_4(z) = \sum_{n \ge 0} c_{H_4}(n) e^{\pi i n z}
\end{equation}
for some $c_{H_4}(n) \in \R$ with $c_{H_4}(0) = 1$ and $c_{H_4}(n) = O(n^k)$ for some $k \in \N$.
Especially, $H_4$ is not cuspidal.
\end{proposition}
\begin{proof}
    \leanok
\end{proof}


We also have a nontrivial relation between these theta functions.
\begin{lemma}\label{lemma:jacobi-identity}\uses{lemma:theta-modular, cor:dim-mf}\lean{jacobi_identity}\leanok
These three theta functions satisfy the \emph{Jacobi identity}
\begin{equation}\label{eqn:jacobi-identity}
H_{2} + H_{4} = H_{3} \Leftrightarrow \Theta_{2}^4 + \Theta_{4}^4 = \Theta_{3}^4.
\end{equation}
\end{lemma}
\begin{proof}\leanok
    Let $f = (H_2 + H_4 - H_3)^{2}$.
    Obviously, $f$ is a modular form of weight $4$ and level $\Gamma(2)$.
    However, by using the transformation rules of $H_2, H_3, H_4$, one have
    \begin{align}
        f|_{S} &= (-H_4 - H_2 + H_3)^{2} = f\\
        f|_{T} &= (-H_2 + H_3 - H_4)^{2} = f
    \end{align}
    so $f$ is actually a modular form of level $1$.
    By considering the limit as $z \to i\infty$, $f$ is a cusp form, so we get $f = 0$ from \eqref{eqn:dimS4}.
\end{proof}

These are also related to $E_4$, $E_6$, and $\Delta$ as follows.
\begin{lemma}\label{lemma:lv1-lv2-identities}\uses{lemma:theta-transform-S-T, lemma:theta-modular, lemma:disc-cuspform}\lean{SpherePacking.ModularForms.E₄_eq_thetaE4,SpherePacking.ModularForms.E₆_eq_thetaE6,Delta_eq_H₂_H₃_H₄}\leanok
We have
\begin{align}
    E_4 &= \frac{1}{2}(H_{2}^{2} + H_{3}^{2} + H_{4}^{2}) = H_{2}^{2} + H_{2}H_{4} + H_{4}^{2} \label{eqn:e4theta} \\
    E_6 &= \frac{1}{2} (H_{2} + H_{3})(H_{3} + H_{4}) (H_{4} - H_{2}) = \frac{1}{2}(H_2 + 2H_4)(2H_2 + H_4)(H_4 - H_2) \label{eqn:e6theta} \\
    \Delta &= \frac{1}{256} (H_{2}H_{3}H_{4})^2. \label{eqn:disctheta}
\end{align}
\end{lemma}
\begin{proof}\leanok
We can prove these similarly as Lemma \ref{lemma:jacobi-identity}.
Right hand sides of \eqref{eqn:e4theta}, \eqref{eqn:e6theta}, and \eqref{eqn:disctheta} are all modular forms of level $\Gamma_1$ and desired weights, where \eqref{eqn:disctheta} is a cusp form since $H_2$ is.
Now the identities follow from the dimension calculations $\dim M_4(\Gamma_1) = \dim M_6(\Gamma_1) = \dim S_{12}(\Gamma_1) = 1$ and comparing the first nonzero $q$-coefficients.
\end{proof}

The \emph{strict} positivity of Jacobi theta functions might needed later.
\begin{corollary}\label{cor:theta-pos}\uses{lemma:jacobi-identity, lemma:theta-transform-S-T}\lean{H₂_imag_axis_pos,H₄_imag_axis_pos,jacobi_identity}\leanok
All three functions $t \mapsto H_2(it), H_3(it), H_4(it)$ are positive for $t > 0$.
\end{corollary}
\begin{proof}\leanok
By Lemma \ref{lemma:jacobi-identity} and the transformation law \eqref{eqn:H2-transform-S}, it is enough to prove the positivity for $\Theta_2(it)$, which is clear from its definition:
\begin{equation}
    \Theta_{2}(it) = \sum_{n \in \Z} e^{- \pi (n + \frac{1}{2})^{2} t} > 0.
\end{equation}
\end{proof}

\subsection{Quasimodular forms and derivatives}

Morally, quasimodular forms can be thought as \emph{modular forms with differentiations}.
It can be defined formally as follows:
Let $f: \mathfrak{H} \to \C$ be a holomorphic function, and let $k$ and $s \ge 0$ be integers.
The function $f$ is a \emph{quasimodular form of weight $k$, level $\Gamma$, and depth $s$} if there exist holomorphic functions $f_0, \dots, f_s : \mathfrak{H} \to \C$ such that
\begin{equation}\label{eqn:quasimod-def}
    (f|_{k}\gamma)(z) = (cz + d)^{-k} f\left(\frac{az + b}{cz + d}\right) = \sum_{j=0}^{s} f_j(z) \left(\frac{c}{cz + d}\right)^j
\end{equation}
for all $z \in \mathfrak{H}$ and $\gamma = \left(\begin{smallmatrix} a&b\\c&d \end{smallmatrix}\right) \in \Gamma$.

By taking $\gamma = \left(\begin{smallmatrix} 1 & 0 \\ 0 & 1 \end{smallmatrix}\right)$, one can check that we should have $f_0 = f$. Thus, a quasimodular form of depth $0$ is just a modular form of same weight and level.
Also, it is easy to see that the space of quasimodular forms is closed under the normalized derivative.

In the Lean formalization, we do not use the above definition of quasimodular forms directly.
Instead, we only use $E_2$ (defined in \Cref{def:E2}) and define the normalized derivative operator $D$ (\Cref{def:derivative}) and the Serre derivative $\partial_k$ (\Cref{def:serre-der}) as the main tools to work with quasimodular forms.

\begin{definition}\label{def:derivative} \lean{D} \leanok
Let $F$ be a quasimodular form.
We define the (normalized) derivative of $F$ as
\begin{equation}\label{eqn:derivative}
    F' = DF := \frac{1}{2\pi i} \frac{\dd}{\dd z} F.
\end{equation}
\end{definition}

Normalizing the derivative by $1/(2 \pi i)$ as in \eqref{eqn:derivative} is standard in the theory of modular forms, since it makes the $q$-expansion coefficients nicer.
\begin{lemma}\label{lemma:der-q-series}\uses{def:derivative}\lean{D_qexp_tsum_pnat}\leanok
We have an equality of operators $D = q \frac{\dd}{\dd q}$.
In particular, the $q$-series of the derivative of a quasimodular form $F(z) = \sum_{n \ge n_0} a_n q^n$ is $F'(z) = \sum_{n \ge n_0} n a_n q^n$.
\end{lemma}
\begin{proof}\leanok
Directly follows from the definition \eqref{def:derivative}, where $\frac{1}{2 \pi i}\frac{\dd}{\dd z}e^{2\pi i n z} = n e^{2\pi i n z}$.
\end{proof}

The most important example of quasimodular form is the weight 2 Eisenstein series $E_2$ (\Cref{def:E2}).
Using it, we can define the \emph{Serre derivative} of a quasimodular form.

\begin{definition}\label{def:serre-der}\uses{def:derivative, def:E2}\leanok \lean{serre_D}
For $k \in \mathbb{R}$, define the weight $k$ Serre derivative $\partial_{k}$ of a modular form $F$ as
\begin{equation}\label{eqn:serre-der}
    \partial_{k}F := F' - \frac{k}{12} E_2 F.
\end{equation}
\end{definition}

\begin{theorem}
\label{thm:serre-der-equiv-action}\uses{def:serre-der, def:E2, lemma:E2-transform-general}\lean{serre_D_slash_equivariant}\leanok
Serre derivative $\partial_{k}$ is equivariant with the slash action of $\mathrm{SL}_{2}(\mathbb{Z})$ in the following sense:
\begin{equation}
    \partial_{k} (F|_{k}\gamma) = (\partial_{k} F)|_{k+2}\gamma, \quad \forall \gamma \in \mathrm{SL}_{2}(\mathbb{Z}).
\end{equation}
\end{theorem}
\begin{proof}\leanok
Let $G = \partial_{k}F = F' - \frac{k}{12}E_2 F$.
From $F \in M_k(\Gamma)$, we have
\begin{equation}
    (F|_{k}\gamma)(z) := (cz + d)^{-k} F\left(\frac{az + b}{cz + d}\right), \quad \gamma = \begin{pmatrix}a & b \\ c & d\end{pmatrix} \in \Gamma.
\end{equation}
By taking the derivative of the above equation, we get
\begin{align}
    \frac{\dd}{\dd z}(F|_{k} \gamma)(z) &= -kc (cz + d)^{-k - 1} F\left(\frac{az + b}{cz + d}\right) + (cz + d)^{-k} (cz + d)^{-2} \frac{\dd F}{\dd z}\left(\frac{az + b}{cz + d}\right) \\
    &= -kc (cz + d)^{-k - 1} F\left(\frac{az + b}{cz + d}\right) + (cz + d)^{-k - 2} \frac{\dd F}{\dd z}\left(\frac{az + b}{cz + d}\right) \\
    &= -kc (cz + d)^{-k - 1} F\left(\frac{az + b}{cz + d}\right) + 2 \pi i (cz + d)^{-k - 2} F'\left(\frac{az + b}{cz + d}\right) \\
    \Leftrightarrow (F|_{k} \gamma)'(z) &= -\frac{kc}{2 \pi i} (cz + d)^{-k - 1} F\left(\frac{az + b}{cz + d}\right) + (cz + d)^{-k - 2} F'\left(\frac{az + b}{cz + d}\right).
\end{align}
Combined with \eqref{eqn:E2-transform-general}, we get
\begin{align}
    ((\partial_k F)|_{k+2}\gamma)(z) &= (cz + d)^{-k-2} \left(F'\left(\frac{az + b}{cz + d}\right) - \frac{k}{12}E_2\left(\frac{az + b}{cz + d}\right)F\left(\frac{az + b}{cz + d}\right)\right) \\
    &= (cz + d)^{-k-2} F'\left(\frac{az + b}{cz + d}\right) - \frac{k}{12} \left(E_2(z) - \frac{6ic}{\pi(cz + d)}\right) \cdot (cz + d)^{-k} F\left(\frac{az + b}{cz + d}\right) \\
    &= (F|_{k}\gamma)'(z) - \frac{k}{12} E_2(z) (F|_{k}\gamma)(z) \\
    &= \partial_{k} (F|_{k}\gamma)(z).
\end{align}
\end{proof}

As a direct consequence of Theorem \ref{thm:serre-der-equiv-action}, we can check that the Serre derivative preserves the modularity of a modular form.
\begin{theorem}\label{thm:serre-der-modularity}\uses{def:serre-der, def:holomorphic-modular-form, thm:serre-der-equiv-action}\lean{serre_D_slash_invariant}\leanok
Let $F$ be a modular form of weight $k$ and level $\Gamma$.
Then, $\partial_{k}F$ is a modular form of weight $k + 2$ of the same level.
\end{theorem}
\begin{proof}\leanok
    Immediate from Theorem \ref{thm:serre-der-equiv-action} since $F|_k\gamma = F$ for all $\gamma \in \Gamma$.
\end{proof}

\begin{remark}
More generally, the following theorem holds: if $F$ is a quasimodular form of weight $k$ and depth $s$, then $\partial_{k-s}F$ is a quasimodular form of weight $k + 2$ \emph{and depth $\le s$} of the same level. We will not prove this here.
\end{remark}

\begin{theorem}\label{thm:ramanujan-formula}\uses{thm:serre-der-modularity, def:serre-der, lemma:E2-transform-general, cor:dim-mf} \lean{ramanujan_E₂, ramanujan_E₄, ramanujan_E₆, ramanujan_E₂', ramanujan_E₄', ramanujan_E₆'} \leanok
We have
\begin{align}
    E_2' &= \frac{E_2^2 - E_4}{12} \label{eqn:DE2} \\
    E_4' &= \frac{E_2 E_4 - E_6}{3} \label{eqn:DE4} \\
    E_6' &= \frac{E_2 E_6 - E_4^2}{2} \label{eqn:DE6}
\end{align}
\end{theorem}
\begin{proof}\leanok
In terms of Serre derivatives, these are equivalent to
\begin{align}
    \partial_{1}E_2 &= -\frac{1}{12} E_4 \label{eqn:SE2} \\
    \partial_{4}E_4 &= -\frac{1}{3} E_6 \label{eqn:SE4} \\
    \partial_{6}E_6 &= -\frac{1}{2} E_4^2 \label{eqn:SE6}
\end{align}
By Theorem \ref{thm:serre-der-modularity}, all the Serre derivatives are, in fact, modular.
To be precise, the modularity of $\partial_{4} E_4$ and $\partial_6 E_6$ directly follows from Theorem \ref{thm:serre-der-modularity}, and that of $\partial_{1}E_2$ follows from \eqref{eqn:E2-transform-general}.
Differentiating and squaring then gives us the following:
\begin{align}
    E_2'|_{4}\gamma &= E_2' - \frac{ic}{\pi(cz + d)} E_2 - \frac{3c^2}{\pi^2 (cz + d)^2} \label{eqn:DE2-transform} \\
    E_2^2|_{4}\gamma &= E_2^2 - \frac{12ic}{\pi(cz + d)} E_2 - \frac{36c^2}{\pi^2 (cz + d)^2} \label{eqn:E2sq-transform}
\end{align}
Hence, \eqref{eqn:DE2}$-\frac{1}{12}$\eqref{eqn:E2sq-transform} is a modular form of weight 4.
By \Cref{cor:dim-mf}, they should be multiples of $E_4, E_6, E_4^2$, and the proportionality constants can be determined by observing the constant terms of $q$-expansions.
\end{proof}
\begin{corollary}\label{cor:logder-disc-E2}\uses{thm:ramanujan-formula, def:disc-definition}\lean{D_Delta_eq_E₂_mul_Delta}\leanok
\begin{equation}\label{eqn:logder-disc-E2}
    \Delta' = E_2 \Delta.
\end{equation}
\end{corollary}
\begin{proof}\leanok
By Ramanujan's formula \eqref{eqn:DE4} and \eqref{eqn:DE6},
\begin{equation}
\Delta' = \frac{3 E_4^2 E_4' - 2 E_6 E_6'}{1728} = \frac{1}{1728} \left(3 E_4^2 \cdot \frac{E_2 E_4 - E_6}{3} - 2 E_6 \cdot \frac{E_2 E_6 - E_4^2}{2}\right) = \frac{E_2(E_4^3 - E_6^2)}{1728} = E_2\Delta.
\end{equation}
\end{proof}

Similar argument allow us to compute (Serre) derivatives of $H_2, H_3, H_4$.
\begin{proposition}\label{prop:theta-der}\uses{def:serre-der, lemma:theta-transform-S-T, lemma:jacobi-identity}\lean{serre_D_two_H₂,serre_D_two_H₃,serre_D_two_H₄}\leanok
We have
\begin{align}
    H_2' &= \frac{1}{6} (H_{2}^{2} + 2 H_{2} H_{4} + E_2 H_2) \label{eqn:H2-der}\\
    H_3' &= \frac{1}{6} (H_{2}^{2} - H_{4}^{2} + E_2 H_3) \label{eqn:H3-der}\\
    H_4' &= -\frac{1}{6} (2H_{2} H_{4} + H_{4}^{2} - E_2 H_4) \label{eqn:H4-der}
\end{align}
or equivalently,
\begin{align}
    \partial_{2} H_{2} &= \frac{1}{6} (H_{2}^{2} + 2 H_{2} H_{4}) \label{eqn:H2-serre-der} \\
    \partial_{2} H_{3} &= \frac{1}{6} (H_{2}^{2} - H_{4}^{2}) \label{eqn:H3-serre-der} \\
    \partial_{2} H_{4} &= -\frac{1}{6} (2H_{2} H_{4} + H_{4}^{2}) \label{eqn:H4-serre-der}
\end{align}
\end{proposition}
\begin{proof}\leanok
Equivalences are obvious from the definition of the Serre derivative.
Define $f_{2}, f_{3}, f_{4}$ be the differences of the left and right hand sides of \eqref{eqn:H2-serre-der}, \eqref{eqn:H3-serre-der}, \eqref{eqn:H4-serre-der}.
\begin{align}
    f_{2} &:= \partial_{2} H_{2} - \frac{1}{6} H_{2}(H_{2} + 2H_{4}) \\
    f_{3} &:= \partial_{2} H_{3} - \frac{1}{6} (H_{2}^2 - H_{4}^2) \\
    f_{4} &:= \partial_{2} H_{4} + \frac{1}{6} H_{4}(2H_{2} + H_{4}).
\end{align}
Then these are a priori modular forms of weight $4$ and level $\Gamma(2)$, and our goal is to prove that they are actually zeros.
By Jacobi's identity \eqref{eqn:jacobi-identity}, we have $f_{2} + f_{4} = f_{3}$.
Also, the transformation rules of $H_2, H_3, H_4$ give
\begin{align}
    f_{2}|_{S} &= -f_{4} \\
    f_{2}|_{T} &= -f_{2} \\
    f_{4}|_{S} &= -f_{2} \\
    f_{4}|_{T} &= f_{3} = f_{2} + f_{4}.
\end{align}
Now, define
\begin{align}
    g &:= (2 H_2 + H_4) f_2 + (H_2 + 2 H_4) f_4 \\
    h &:= f_{2}^{2} + f_{2}f_{4} + f_{4}^{2}.
\end{align}
Then one can check that both $g$ and $h$ are invariant under the actions of $S$ and $T$, hence they are modular forms of level $1$.
Also, by analyzing the limit of $g$ and $h$ as $z \to i \infty$, one can see that $g$ and $h$ are cusp forms, hence $g = h = 0$ by \eqref{eqn:dimS6} and \eqref{eqn:dimS8}.
This implies
\begin{align}
    3 E_4 f_2^{2} &= 3 (H_2^2 + H_2 H_4 + H_4^2) f_2^{2} = ((2 H_2 + H_4)^{2} - (2H_2 + H_4)(H_2 + 2H_4) + (H_2 + 2H_4)^{2}) f_2^{2}\\
    &= (2 H_2 + H_4)^{2} (f_2^2 + f_2 f_4 + f_4^2) = 0
\end{align}
and by considering $q$-series ($E_4$ has an invertible $q$-series), we get $f_2 = 0$.
\end{proof}

\begin{theorem}\label{thm:serre-der-prod-rule}\uses{def:serre-der} \lean{serre_D_mul} \leanok
The Serre derivative satisfies the following product rule: for any quasimodular forms $F$ and $G$,
\begin{equation}
    \partial_{w_1 + w_2} (FG) = (\partial_{w_1}F)G + F (\partial_{w_2}G).
\end{equation}
\end{theorem}
\begin{proof}\leanok
    It follows from the definition:
    \begin{align}
        \partial_{w_1 + w_2} (FG) &= (FG)' - \frac{w_1 + w_2}{12} E_2 (FG) \\
        &= F'G + FG' - \frac{w_1 + w_2}{12} E_2(FG) \\
        &= \left(F' - \frac{w_1}{12}E_2 F\right)G + F \left(G' - \frac{w_2}{12}E_2 G\right) \\
        &= (\partial_{w_1}F)G + F(\partial_{w_2}G).
    \end{align}
\end{proof}

We also have the following useful theorem for proving positivity of quasimodular forms on the imaginary axis, which is \cite[Proposition 3.5, Corollary 3.6]{Lee}.
\begin{theorem}\label{thm:anti-serre-der-pos}\uses{def:serre-der, cor:logder-disc-E2}\lean{antiSerreDerPos}\leanok
Let $F$ be a holomorphic quasimodular cusp form with real Fourier coefficients.
Assume that there exists $k$ such that $(\partial_{k}F)(it) > 0$ for all $t > 0$.
If the first Fourier coefficient of $F$ is positive, then $F(it) > 0$ for all $t > 0$.
\end{theorem}
\begin{proof}\leanok
By \eqref{eqn:logder-disc-E2}, we have
\begin{align}
    \frac{\dd}{\dd t} \left( \frac{F(it)}{\Delta(it)^{\frac{k}{12}}}\right)
    &= (-2 \pi) \frac{F'(it) \Delta(it)^{\frac{k}{12}} - F(it) \frac{k}{12} E_{2}(it) \Delta(it)^{\frac{k}{12}}}{\Delta(it)^{\frac{k}{6}}} \\
    &= (-2 \pi) \frac{(\partial_{k} F)(it)}{\Delta(it)^{\frac{k}{12}}}  < 0,
\end{align}
hence
\[
t \mapsto \frac{F(it)}{\Delta(it)^{\frac{k}{12}}}
\]
is monotone decreasing.
Because of the assumption on the positivity of the first nonzero Fourier coefficient of $F$, $F(it) > 0$ for sufficiently large $t$ since
\[
F = \sum_{n \geq n_{0}} a_{n} q^{n} \Rightarrow e^{2 \pi n_{0} t} F(it) = a_{n_{0}} + e^{-2 \pi t}\sum_{n\geq n_{0} + 1} a_{n} e^{-2 \pi (n - n_{0} - 1)t}
\]
and $\lim_{t \to \infty} e^{2 \pi n_{0}t} F(it) = a_{n_0} > 0$, hence the result follows.
\end{proof}
