The definition of density given in \cref{sec:sphere-packings} is inconvenient to work with, especially when our packing is a structured one, such as a periodic/lattice packing. This section fixes this problem.

All statements and proofs in this section are formalized in Lean.

Observe that the finite density evaluated at some $R > 0$ measures the density of sphere packings within a bounded, open ball of radius $R$. As one might expect, there is a relationship between the finite density and the number of centers in the ball of radius $R$.

\begin{lemma}\label{lemma:sp-finite-density-bound}\lean{SpherePacking.finiteDensity_le,SpherePacking.finiteDensity_ge}\uses{def:SpherePacking.finiteDensity}\leanok
  For any $R > 0$,
  \[
    \left|X \cap \mathcal{B}_d\left(R - \frac{r}{2}\right)\right| \cdot \frac{\mathrm{Vol}\left(\mathcal{B}_d\left(\frac{r}{2}\right)\right)}{\mathrm{Vol}(\mathcal{B}_d(R))}
    \leq \Delta_{\mathcal{P}}(R)
    \leq \left|X \cap \mathcal{B}_d\left(R + \frac{r}{2}\right)\right| \cdot \frac{\mathrm{Vol}\left(\mathcal{B}_d\left(\frac{r}{2}\right)\right)}{\mathrm{Vol}(\mathcal{B}_d(R))}
  \]
\end{lemma}
\begin{proof}\uses{def:SpherePacking}\leanok
  The high level idea is to prove that $\mathcal{P} \cap \mathcal{B}_d(R) = \left(\bigcup_{x \in X} \mathcal{B}_d\left(x, \frac{r}{2}\right)\right) \subseteq \bigcup_{x \in X \cap \mathcal{B}_d\left(R + \frac{r}{2}\right)} \mathcal{B}_d\left(x, \frac{r}{2}\right)$, and a similar inequality for the upper bound. The rest follows by rearranging and using the fact that the balls are pairwise disjoint.
\end{proof}

Suppose further that $X$ is a periodic packing w.r.t. the lattice $\Lambda \subseteq \R^d$. Let $\mathcal{D}$ be a (bounded) fundamental region of $\Lambda$, say the parallelopiped $[0, 1]^n\Lambda$, and let $L$ be the bound on the norm of vectors in $\mathcal{D}$, i.e. a number satisfying $\forall x \in \mathcal{D}, \|x\| \leq L$.

\begin{lemma}\label{lemma:lattice-points-bound}\lean{PeriodicSpherePacking.aux2_ge',PeriodicSpherePacking.aux2_le'}\leanok
  For all $R$, we have the following inequality relating the number of lattice points from $\Lambda$ in a ball with the volume of the ball and the fundamental region $\mathcal{D}$:

  \[
    \frac{\mathrm{Vol}(\mathcal{B}_d(R - L))}{\mathrm{Vol}(\mathcal{D})}
    \leq \left|\Lambda \cap \mathcal{B}_d(R)\right|
    \leq \frac{\mathrm{Vol}(\mathcal{B}_d(R + L))}{\mathrm{Vol}(\mathcal{D})}
  \]
\end{lemma}
\begin{proof}\leanok
For the first inequality, it suffices to prove that $\mathcal{B}_d(R - L) \subseteq \bigcup_{x \in \Lambda \cap \mathcal{B}_d(R)} (x + \mathcal{D})$, since the cosets on the right are disjoint. For a vector $v \in \mathcal{B}_d(R - L)$, we have $\|v\| < R - L$ by definition. Since $\mathcal{D}$ is a fundamental domain, there exists a lattice point $x \in \Lambda$ such that $v \in x + \mathcal{D}$. Rearranging gives $v - x \in \mathcal{D}$, which means $\|v - x\| \leq L$. By the triangle inequality, $\|x\| < R$, i.e. $x \in \mathcal{B}_d(L)$, concluding the proof.

For the second inequality, we prove that $\bigcup_{x \in \Lambda \cap \mathcal{B}_d(R)} (x + \mathcal{D}) \subseteq \mathcal{B}_d(R + L)$. Consider a vector $x \in \Lambda \cap \mathcal{B}_d(R)$ and a vector $y \in x + \mathcal{D}$. From above, we know $\|x\| < R$ and $\|y - x\| \leq L$, so $\|y\| < R + L$, concluding the proof.
\end{proof}

To link \cref{lemma:sp-finite-density-bound} and \cref{lemma:lattice-points-bound}, we need a lemma relating $|\Lambda \cap \mathcal{B}|$ with $|X \cap \mathcal{B}|$, which is what the following lemma does:

\begin{lemma}\label{lemma:periodic-points-bounds}\lean{PeriodicSpherePacking.aux2_ge',PeriodicSpherePacking.aux2_le'}\leanok
  For all $R$, we have the following inequality relating the number of points from $X$ (periodic w.r.t. $\Lambda$) in a ball with the number of points from $\Lambda$:
  \[
    \left|\Lambda \cap \mathcal{B}_d(R - L)\right|\left|X / \Lambda\right|
    \leq \left|X \cap \mathcal{B}_d(R)\right|
    \leq \left|\Lambda \cap \mathcal{B}_d(R + L)\right|\left|X / \Lambda\right|
  \]
\end{lemma}
\begin{proof}\leanok
  For the first inequality, we notice that $\bigcup_{x \in \Lambda \cap \mathcal{B}_d(R - L)} (x + \mathcal{D}) \subseteq \mathcal{B}_d(R)$, because for $x \in \Lambda \cap \mathcal{B}_d(R - L)$ and $y \in x + \mathcal{D}$, we have $\|x\| < R - L$ and $\|y - x\| \leq L$, so $\|y\| < R$ by triangle inequality. Intersecting both sides with $X$ and simplifying, we have

  \[
    \left(\bigcup_{x \in \Lambda \cap \mathcal{B}_d(R - L)} (x + \mathcal{D})\right) \cap X = \bigcup_{x \in \Lambda \cap \mathcal{B}_d(R - L)} ((x + \mathcal{D}) \cap X) \subseteq \mathcal{B}_d(R) \cap X
  \]

  Consider the (finite) cardinality on both sides and noting that $|(x + \mathcal{D}) \cap X| = |X / \Lambda|$ for all $x$, we see that $|\Lambda \cap \mathcal{B}_d(R - L)||X / \Lambda| \leq |X \cap \mathcal{B}_d(R)|$, as desired.

  The proof of the second inequality is similar. We again observe that $\mathcal{B}_d(R) \subseteq \bigcup_{x \in \Lambda \cap \mathcal{B}_d(R + L)} (x + \mathcal{D})$, which follows from the tiling property of fundamental domain (i.e. every point can be translated by a $\Lambda$ lattice point into $\mathcal{D}$). Intersecting both sides with $X$ and considering cardinality of both sides concludes the proof.

  There are several technicalities when formalising in Lean, such as having to prove $|\Lambda \cap \mathcal{B}_d(R)|$ is countable and finite. Those are handled at \texttt{aux3}.
\end{proof}

When we combine the inequalities above, we need one additional computational lemma.

\begin{lemma}\label{lemma:volume-ball-ratio-limit}\lean{volume_ball_ratio_tendsto_nhds_one''}\leanok
  For any constant $C > 0$, we have

  \[
    \lim_{R \to \infty} \frac{\mathrm{Vol}(\mathcal{B}_d(R))}{\mathrm{Vol}(\mathcal{B}_d(R + C))} = 1
  \]
\end{lemma}
\begin{proof}\leanok
  Write out the formula for volume of a ball and simplify. More specifically, we have $\mathrm{Vol}(\mathcal{B}_d(R)) = R^d \pi^{d / 2} / \Gamma\left(\frac{d}{2} + 1\right)$, so $\mathrm{Vol}(\mathcal{B}_d(R)) / \mathrm{Vol}(\mathcal{B}_d(R + C)) = R^d / (R + C)^d = \left(1 - \frac{1}{R + C}\right)^d = 1$.
\end{proof}

Finally, we can relate the density of a periodic sphere packing to the natural definition of density given by any of its fundamental domain:

\begin{theorem}\label{theorem:psp-density}\uses{def:SpherePacking.density}\lean{PeriodicSpherePacking.density_eq}\leanok
  For a periodic sphere packing $\mathcal{P} = \mathcal{P}(X)$ with centers $X$ periodic to the lattice $\Lambda$ and separation $r$,

  \[
    \Delta_{\mathcal{P}} = |X / \Lambda| \cdot \frac{\Vol{\mathcal{B}_d(r / 2)}}{\Vol{\R^d / \Lambda}}
  \]
\end{theorem}
\begin{proof}\uses{lemma:sp-finite-density-bound,lemma:lattice-points-bound,lemma:periodic-points-bounds,lemma:volume-ball-ratio-limit}\leanok
  Fix any fundamental domain $\mathcal{D}$ (induced by any basis) of the lattice $\Lambda$. Combining \cref{lemma:sp-finite-density-bound}, \cref{lemma:lattice-points-bound} and \cref{lemma:periodic-points-bounds}, we get the following inequality for the \textit{finite} density:

  \[
    |X / \Lambda| \cdot \frac{\Vol{\mathcal{B}_d(r / 2)}}{\Vol{\R^d / \Lambda}} \cdot \frac{\Vol{\mathcal{B}_d(R - r / 2 - 2L)}}{\Vol{\mathcal{B}_d(R)}}
    \leq \Delta_{\mathcal{P}}(R)
    \leq |X / \Lambda| \cdot \frac{\Vol{\mathcal{B}_d(r / 2)}}{\Vol{\R^d / \Lambda}} \cdot \frac{\Vol{\mathcal{B}_d(R + r / 2 + 2L)}}{\Vol{\mathcal{B}_d(R)}}
  \]

  Taking limit on both sides as $R \to \infty$ and apply the Sandwich theorem and \cref{lemma:volume-ball-ratio-limit}, we get

  \[
    \Delta_{\mathcal{P}} = \limsup_{R \to \infty} \Delta_{\mathcal{P}}(R) = \lim_{R \to \infty} \Delta_{\mathcal{P}}(R) = |X / \Lambda| \cdot \frac{\Vol{\mathcal{B}_d(r / 2)}}{\Vol{\R^d / \Lambda}}
  \]
\end{proof}
